---
title: "DS 3001 Final Project"
author: "Jess Laudie"
date: "12/1/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(ggplot2)
```

# Question and Background 


# Data Processing 
```{r}
song_metrics <- read.csv("C:/Users/Owner/OneDrive/Desktop/DS-Final-Project/TaylorSwift.csv")

```


# Exploraory Analysis - Song Metrics and Spotify Features 

Some key Spotify metrics that we can explore when discussing Taylor Swift's music is danceability, valence, energy, and length. 

- Danceability: This is how suitable a track is for dancing based on factors like tempo and rhythm. 0.0 is the least danceable and 1.0 is the most danceable.
- Valence: This describes the positiveness displayed in each track. It is scored from 0.0 to 1.0 with 1.0 being the least positive. Tracks with high valence sounds are likely to be more happy, cheerful, euphoric, etc. while tracks with low valence have more negative sentiments so the music overall sounds more sad, depressing, mad, etc.
- Energy: A measure from 0.0 to 1.0 and represents the intensity and activity found in each track. A measure of 1.0 feels very fast, loud, and noisy, while a meausure of 0 is the opposite. 
- Length: duration of the track in milliseconds.


```{r}
# Danceability
song_metrics = song_metrics[-c(1,6:12,14:15)]
View(song_metrics)
summary(song_metrics)
# average danceability = 0.5925
# average valence = 0.4173
# average energy = 0.5777
# average length = 235662 milliseconds or 3.9 minutes

# Song length for each album 
ggplot(data = song_metrics, mapping = aes(x = Album, y = Length)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 5, color = "pink")

# Song danceability for each album 
ggplot(data = song_metrics, mapping = aes(x = Album, y = danceability)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 5, color = "tomato")

# Song energy for each album 
ggplot(data = song_metrics, mapping = aes(x = Album, y = energy)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 5, color = "blue")

# Song valence for each album 
ggplot(data = song_metrics, mapping = aes(x = Album, y = valence)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 5, color = "yellow")

# Relationships between variables

ggplot(data = song_metrics, aes(x = valence, y = Length)) +
  geom_point(alpha = 2, aes(color = Album))
# More positive songs that rank higher in valence are the shortest songs and
#Longest songs rank the lowest in valence so they have more sentiments that relate more to sadness, anger, etc.
# Good mix of songs from each album in terms of theri valence 



```


```{r}

```


# Sentiment Analysis

Taylor Swift 

Fearless 


Speak Now 

```{r}

```

Red

```{r}

```

1989
Reputation
Lover
Folklore 
Evermore
Fearless TV
Red TV

# Conclusion 

#Future Work 
